### 2014-10-30 intro.tex
Formatting operational semantics
edits

\begin{itemize}
\setlength{\itemsep}{2pt}
\item We introduce \name, a language for declaratively describing eventually
	consistent programs that manipulate replicated data types. Contracts are used
	to specify fine-grained application-level consistency properties, and are
	analyzed to assign the most efficient and sound store consistency level to
	the corresponding operation.
\item \name supports coordination-free transactions over arbitrary datatypes.
	We extend our contract language to express fine-grained transaction isolation
	guarantees, and utilize the contract enforcement system to automatically
	assign the correct and most efficient isolation level for a transaction.
\item We provide metatheory that certifies the soundness of our contract enforcement
	system, and ensures that an operation is only executed if the required
	conditions on consistency are met.
\item \name is implemented as a transparent shim layer over
  Cassandra~\cite{Cassandra}, a well-known general-purpose data store.  An
  extensive experimental evaluation over a collection of real-world
  applications, including a Twitter-like micro-blogging site and an
  eBay-like auction site illustrates the practicality of our approach.
\end{itemize}

>>>>

\begin{itemize}
\setlength{\itemsep}{2pt}
\item We introduce \name, a Haskell library that supports the description
  and validation of replicated data types found in programs intended to
  execute on eventually consistent stores. Contracts are used to specify
  fine-grained application-level consistency properties, and are analyzed to
  assign the most efficient and sound store consistency level to the
  corresponding operation.
\item \name supports coordination-free transactions over arbitrary datatypes.
	We extend our contract language to express fine-grained transaction isolation
	guarantees, and utilize the contract enforcement system to automatically
	assign the correct and most efficient isolation level for a transaction.
\item We provide metatheory that certifies the soundness of our contract
	enforcement system, and ensures that an operation is only executed if the
	required conditions on consistency are met.
\item An implementation of \name as a transparent shim layer over
	Cassandra~\cite{Cassandra}, a well-known general-purpose data store.  An
	extensive experimental evaluation over a collection of real-world
	applications, including a Twitter-like micro-blogging site and an eBay-like
	auction site illustrates the practicality of our approach.
\end{itemize}

### 2014-10-30 motivation.tex
Edits
edits

\noindent An operation takes a list of effects (the \emph{context} for the operation), an
argument and returns a result along with an optional effect. Read-only
operations do not return any effect. If an effect is returned, it is added to
the state of the object at the current replica, and asynchronously sent to
other replicas. The implementation of the bank account operations in \name is
given below:

\begin{codehaskell}
data Acc = Deposit_ Int | Withdraw_ Int
				 | GetBalance_

@@

mkOperName ''Acc
\end{codehaskell}

The datatype \cf{Acc} represents the effect type for the bank account. The
context of the operations is a snapshot of the state of the object at some
replica. In this sense, every operation on the RDT is atomic, and thus
permitting sequential reasoning for implementing eventually consistent data
types. We have implemented a large corpus of RDTs for realistic benchmarks
including shopping carts, auction and micro-blogging sites in few tens of lines
of code.

The last line in the above code snippet uses Template Haskell~\cite{} to
generate the following datatype declaration:

@@

\name helps facilitate the mapping of operations to appropriate consistency
levels by letting the programmer declare application-level consistency
constraints as \emph{contracts} that axiomatically specify the set of allowed
executions involving this operation.  In the case of the bank account example,
any execution that does not exhibit the anomalies described in the previous
section is a \emph{well-formed} execution on the bank account object.  By
specifying the set of legal executions for each data type in terms of a trace
of operation invocations on that type, \name\ ensures that all executions over
that type are well-formed.

@@

\name contract language is presented in Figure~\ref{code:ctrt_lang}. The
language includes primitive relations \cf{Vis},\cf{So} and \cf{SameObj},
operators to compose relations ($\cup,\cap$ and \cf{transClosure}), and check
inclusion ($\in$). The language also includes standard propositional logic
operators, and a prenex quantified $\forall$ operator. The functions \cf{vis},
\cf{so} and \cf{sameObj} are derived entities. For example, \cf{vis a b = (a,b)
$\in$ Vis}.

In our running example, it is clear that in order to preserve the integrity
constraint, the \cf{withdraw} operation must be strongly consistent. That is,
given two \cf{withdraw} operations $a$ and $b$, either $a$ is visible to $b$ or
vice versa. We express this application-level consistency as the following
contract:

@@

\noindent For \cf{getBalance}, we construct the following contract:

\begin{codehaskell}
getBalanceCtrt x = (forall) [Deposit, Withdraw] $ \a ->
	(forall) [Deposit] $ \b -> (forall) [Withdraw] $ \c ->
		lift $ ((a,x) (- (So cap SameObj) => vis a x) wedge
  				 (vis b c wedge vis c x => vis b x)
\end{codehaskell}

\noindent Here, \cf{a} is an effect emitted by either a \cf{deposit} or a
\cf{withdraw} operation. The fourth line in the above snippet says that if a
\cf{deposit} or a \cf{withdraw} operation precedes a \cf{getBalance} operation
in session order, and is applied on the same object as the \cf{getBalance}
operation, then it must be the case that the \cf{getBalance} operation
witnesses the effects of that preceding operations. The last line says that if
a \cf{withdraw} $c$ is visible to \cf{getBalance}, then all \cf{deposit}
operations $b$ visible to the \cf{withdraw} should also be visible to
\cf{getBalance}. This prevents negative balance anomalies.

@@

\begin{codehaskell}
data Quelea a
instance Monad Quelea
invoke :: (Arg a, Res r)
			 => Key -> OperName -> a -> Quelea r
\end{codehaskell}

>>>>

\noindent An operation takes a list of effects (the \emph{context} for the operation), an
argument and returns a result along with an optional effect. Read-only
operations do not return any effect. If an effect is returned, it is added to
the state of the object at the current replica, and asynchronously sent to
other replicas. The implementation of these operations is
given in Figure~\ref{fig:ex}.

\begin{figure}[htb]
\begin{codehaskell}
data Acc = Deposit_ Int | Withdraw_ Int
				 | GetBalance_

@@

mkOperName ''Acc
\end{codehaskell}
\caption{Definition of a replicated datatype in Quelea written in Haskell syntax.}
\label{fig:ex}
\end{figure}

The datatype \cf{Acc} represents the effect type for the bank account. The
context of the operations is a snapshot of the state of the object at some
replica. In this sense, every operation on the RDT is atomic, and thus
permits sequential reasoning for implementing eventually consistent data
types. We have implemented a large corpus of RDTs for realistic benchmarks
including shopping carts, auction and micro-blogging sites, all in few tens
of lines of code.

The last line in the code snippet uses Template Haskell~\cite{} to
generate the following datatype declaration:

@@

\name helps facilitate the mapping of operations to appropriate consistency
levels by letting the programmer additionally declare application-level consistency
constraints as \emph{contracts} that axiomatically specify the set of allowed
executions involving this operation.  In the case of the bank account example,
any execution that does not exhibit the anomalies described in the previous
section is a \emph{well-formed} execution on the bank account object.  By
specifying the set of legal executions for each data type in terms of a trace
of operation invocations on that type, \name\ ensures that all executions over
that type are well-formed.

@@

\name The contract language is presented in Figure~\ref{code:ctrt_lang}. The
language includes primitive relations \cf{Vis},\cf{So} and \cf{SameObj},
operators to compose relations ($\cup,\cap$ and \cf{transClosure}), and check
inclusion ($\in$). The language also includes standard propositional logic
operators, and a prenex quantified $\forall$ operator. The functions \cf{vis},
\cf{so} and \cf{sameObj} are derived entities. For example, \cf{vis a b = (a,b)
$\in$ Vis}.

In our running example, it is clear that in order to preserve the critical
integrity constraint, the \cf{withdraw} operation must be strongly
consistent. That is, given two \cf{withdraw} operations $a$ and $b$, either
$a$ is visible to $b$ or vice versa. We express this application-level
consistency as the following contract:

@@

\noindent For \cf{getBalance}, we construct the following contract:
\begin{codehaskell}
getBalanceCtrt x = (forall) [Deposit, Withdraw] $ \a ->
	(forall) [Deposit] $ \b -> (forall) [Withdraw] $ \c ->
		lift $ ((a,x) (- (So cap SameObj) => vis a x) wedge
  				 (vis b c wedge vis c x => vis b x)
\end{codehaskell}
\noindent Here, \cf{a} is an effect emitted by either a \cf{deposit} or a
\cf{withdraw} operation. The fourth line in the above snippet says that if a
\cf{deposit} or a \cf{withdraw} operation precedes a \cf{getBalance} operation
in session order, and is applied on the same object as the \cf{getBalance}
operation, then it must be the case that the \cf{getBalance} operation
witnesses the effects of that preceding operations. The last line says that if
a \cf{withdraw} $c$ is visible to \cf{getBalance}, then all \cf{deposit}
operations $b$ visible to the \cf{withdraw} should also be visible to
\cf{getBalance}. This prevents negative balance anomalies.

@@

\begin{codehaskell}
data Quelea a
instance Monad Quelea
invoke :: (Table t, Key k, Arg a, Res r)
			 => t -> k -> OperName -> a -> Quelea r
\end{codehaskell}
